---
title: "Calculations draft"
author: "Mitchell Ardolf"
date: "10/2/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Below is what I personally did to calculate our two values. 




9. What two features are you planning to use for your first challenge?

Number of dark pixels in the top half.
Width of the number.

10. Using the MNIST dataset select two cases (from your training dataset) of each of the digits assigned to your group. Calculate the two features from point 9 on each of those cases. Are the two features similar across the two different types of digits? Use the following code to get you started.

```{r}
library(dplyr)
library(dslabs)
library(tidymodels)
conflicted::conflicts_prefer(dplyr::filter)
mnist <- read_mnist("~/Mscs 341 F23/Class/Data")

# Plots a digit as an image
plot_digit <- function(image) {
  a_digit <- matrix(image, nrow=28)
  image(a_digit[,28:1])
}

# This function returns a matrix with all the images corresponding to a digit (0-9) 
# The matrix will have 784 columns and the number of rows corresponds to the
# number of images
get_image_digit <- function (digit) {
  idx <- mnist$train$labels==digit
  images <- mnist$train$images[idx,]
  return (images)
}

get_image_digit_test <- function (digit) {
  idx <- mnist$test$labels==digit
  images <- mnist$test$images[idx,]
  return (images)
}

#sampling time
data1<-as.data.frame(get_image_digit(1))
data4<-as.data.frame(get_image_digit(4))
DataUnsampled<-rbind(data1,data4)
set.seed(897)
SampleVec <- c(1:12584)
SampleRows<-sample(SampleVec, 800)
train1and4<-DataUnsampled%>%
  mutate(rn = row_number())%>%
  mutate(label = ifelse(rn <=6742, 1, 4))%>%
  dplyr::filter(rn %in% SampleRows)


data1test<-as.data.frame(get_image_digit_test(1))
data4test<-as.data.frame(get_image_digit_test(4))
a<-1135+982
DataUnsampledTest<-rbind(data1test,data4test)
SampleVecTest <- c(1:2117)
SampleRowsTest<-sample(SampleVecTest, 200)
test1and4<-DataUnsampledTest%>%
  mutate(rn = row_number())%>%
  mutate(label = ifelse(rn <=1135, 1, 4))%>%
  dplyr::filter(rn %in% SampleRowsTest)


train1and4%>%select(label)
train1and4LABELS <- train1and4$label

test1and4%>%select(label)
test1and4LABELS <- test1and4$label


# dark pixels top half and vector storage
count_dark_pixels_top_half <- function (image){
  image_top_half <- image[1:392]
  dark_idx <- image_top_half>128
  num_pixels <- sum(dark_idx)
  return(num_pixels)
}

#test this
imagetheory <- vector(mode = "integer", length = 784)
for (i in 1:392){
  n<-i  #+392
  imagetheory[n]<-255
  
}

plot_digit(imagetheory)
plot_digit(mnist$train$images[3,])




train_dark_pixels_top_half <- vector(mode="integer", length=800)
for (i in 1:800){
  imagedf<-train1and4 %>%
      select(1:784)%>%
    slice(i)
  imagevec<-as.numeric(as.vector(imagedf))
  train_dark_pixels_top_half[i]<-count_dark_pixels_top_half(image = imagevec)
  
}
test_dark_pixels_top_half <- vector(mode="integer", length=200)
for (i in 1:200){
  imagedf<-test1and4 %>%
      select(1:784)%>%
    slice(i)
  imagevec<-as.numeric(as.vector(imagedf))
  test_dark_pixels_top_half[i]<-count_dark_pixels_top_half(image = imagevec)
  
}

#this was taken from my hw
get_left_border <- function(image_vec){ 
  entered_df<-data.frame(image_vec=image_vec, n=c(1:784))
first_non_zero<-entered_df%>%
  filter(image_vec > 0)%>%
  arrange(n)%>%
  slice(1)
remainder <- first_non_zero$n %% 28
#zero_v <- vector(mode="integer", length=784)
#for (i in 1:28){
#n <- (28-remainder) + (i-1)*28
#  zero_v[n]<-255}
#return(zero_v)
return(remainder)
}

train_leftrem <- vector(mode="integer", length=800)
for (i in 1:800){
  imagedf<-train1and4 %>%
      select(1:784)%>%
    slice(i)
  imagevec<-as.numeric(as.vector(imagedf))
  train_leftrem[i]<-get_left_border(imagevec)
  
}
test_leftrem <- vector(mode="integer", length=200)
for (i in 1:200){
  imagedf<-test1and4 %>%
      select(1:784)%>%
    slice(i)
  imagevec<-as.numeric(as.vector(imagedf))
  test_leftrem[i]<-get_left_border(imagevec)
  
}


get_right_border <- function(image_vec){ 
  a_digit <- matrix(image_vec, nrow=28)
  image_vec_rev <- c(a_digit[28:1,])
  entered_df<-data.frame(image_vec=image_vec_rev, n=c(1:784))
first_non_zero<-entered_df%>%
  filter(image_vec>0)%>%
  arrange(n)%>%
  slice(1)
remainder <- first_non_zero$n %% 28
#zero_v <- vector(mode="integer", length=784)
#for (i in 1:28){
#n <- (28-remainder) + (i-1)*28
 # zero_v[n]<-255}
#return((zero_v))
return(remainder)
}

train_rightrem <- vector(mode="integer", length=800)
for (i in 1:800){
  imagedf<-train1and4 %>%
      select(1:784)%>%
    slice(i)
  imagevec<-as.numeric(as.vector(imagedf))
  train_rightrem[i]<-get_right_border(imagevec)
  
}
test_rightrem <- vector(mode="integer", length=200)
for (i in 1:200){
  imagedf<-test1and4 %>%
      select(1:784)%>%
    slice(i)
  imagevec<-as.numeric(as.vector(imagedf))
  test_rightrem[i]<-get_right_border(imagevec)
  
}

traindf <- data.frame(label = train1and4LABELS, DTH = train_dark_pixels_top_half, right = train_rightrem, left = train_leftrem)%>%
  mutate(width = abs(right - left))
testdf <- data.frame(label = test1and4LABELS, DTH = test_dark_pixels_top_half, right = test_rightrem, left = test_leftrem)%>%
  mutate(width = abs(right - left))


traindf%>%
  ggplot(mapping = aes(x = label, group = label, y = DTH))+
  geom_boxplot()

traindf%>%
  ggplot(mapping = aes(x = label, group = label, y = width))+
  geom_boxplot()
```













Evans Calculations



```{r echo=FALSE}

# Plots a digit as text matrix on the terminal
plot_text <- function(image) {
  a_digit <- matrix(image, nrow=28)
  even_pos <- seq(4,24,2)
  (a_digit[even_pos,even_pos])
}

# Plots a digit as an image
plot_digit <- function(image) {
  a_digit <- matrix(image, nrow=28)
  image(a_digit[,28:1])
}

# This function outputs a contrasted image 
# Notice an image is represented as a vector of 784 number from 0 to 255.
contrast_digit <- function(image){
  dark_idx = (image>128)
  image[dark_idx]=255   # Turn dark pixel indexes to 255
  image[!dark_idx]=0    # Turn non-dark pixel indexes to 0 (we use symbol !)
  return (image)
}

# This function calculates the number of dark pixels from an image
# Notice an image is represented as a vector of 784 number from 0 to 255.
count_dark_pixels <- function (image){
  dark_idx <- image>128
  num_pixels <- sum(dark_idx)
  return(num_pixels)
}

count_light_pixels <- function (image){
  light_idx <- image<128
  num_pixels <- sum(light_idx)
  return(num_pixels)
}

# This function returns a matrix with all the images corresponding to a digit (0-9) 
# from your mnist training dataset.
# The matrix will have 784 columns and the number of rows corresponds to the
# number of images
get_image_train_digit <- function (digit) {
  idx <- mnist$train$labels==digit
  images <- mnist$train$images[idx,]
  return (images)
}

# This function returns a matrix with all the images corresponding to a digit (0-9) 
# from your mnist testing dataset.
# The matrix will have 784 columns and the number of rows corresponds to the
# number of images
get_image_test_digit <- function (digit) {
  idx <- mnist$test$labels==digit
  images <- mnist$test$images[idx,]
  return (images)
}
```

```{r}
library(dslabs)
mnist <- read_mnist("~/Mscs 341 F23/Class/Data")

# Plots a digit as an image
plot_digit <- function(image) {
  a_digit <- matrix(image, nrow=28)
  image(a_digit[,28:1])
}

# This function returns a matrix with all the images corresponding to a digit (0-9) 
# The matrix will have 784 columns and the number of rows corresponds to the
# number of images
get_image_digit_train <- function (digit) {
  idx <- mnist$train$labels==digit
  images <- mnist$train$images[idx,]
  return (images)
}
test <- get_image_digit(4)



get_image_digit_test <- function (digit) {
  idx <- mnist$test$labels==digit
  images <- mnist$test$images[idx,]
  return (images)
}

contrast_digit <- function(image){
  dark_idx = (image>128)
  image[dark_idx]=255   # Turn dark pixel indexes to 255
  image[!dark_idx]=0    # Turn non-dark pixel indexes to 0 (we use symbol !)
  return (image)
}



# Create a sample matrix


```
functions 

```{r}
get_width <- function(image){ 
  #the function takes an image and looks at all columns those columns that contain pixels are counted.
  a_digit <- matrix(image, nrow = 28)
  adigit <- a_digit[, 28:1]
  i <- 1
  j <- 0

  while (i <= 28) {
    if (sum(adigit[i, ]) > 0) {
      j <- j + 1
    }
    
    i <- i + 1
  }
  return(j)
  
}
```


```{r}
plot_digit((mnist$train$images[7,]))



plot_digit(contrast_digit(mnist$train$images[7,]))

get_width(contrast_digit(mnist$train$images[7,]))

count_dark_pixels(mnist$train$images[7, 1:392])
count_dark_pixels(contrast_digit(mnist$train$images[7, 1:392]))

```


```{r}
get_size <- function(images){
  
  n_images <- nrow(images)
  
  # We use a for loop to calculate the darkness of each digit
  size_val <- vector(mode="integer", length=n_images)
  for (i in 1:n_images) {
    size_val[i] = get_width(images[i,]) #calcultate the darkness of each image provided 
  }
  return (size_val)
}


get_size (minist)

get_size_light <- function(images){
  
  n_images <- nrow(images)
  
  # We use a for loop to calculate the darkness of each digit
  size_val <- vector(mode="integer", length=n_images)
  for (i in 1:n_images) {
    size_val[i] = get_width(contrast_digit(images[i,])) #width of the image
  }
  return (size_val)
}




count_dark_pixels <- function (image){ # This function counts the dark pixels in the loop 
  dark_idx <- image>128
  num_pixels <- sum(dark_idx)
  return(num_pixels)
}

get_num_pixels <- function(images){
  # We get all images corresponding to a digit

  n_images <- nrow(images)
  
  # We use a for loop to calculate the ink of each digit
  dark <- vector(mode="integer", length=n_images)
  for (i in 1:n_images) {
    dark[i] = count_dark_pixels(images[i, 1:392])
  }
  return (dark)
}




#get_num_pixels_light <- function(images){
  # We get all images corresponding to a digit

  #n_images <- nrow(images)
  
  # We use a for loop to calculate the ink of each digit
  #dark <- vector(mode="integer", length=n_images)
  #for (i in 1:n_images) {
    #dark[i] = count_dark_pixels(contrast_digit(images[i, 1:392]))
  #}
  #return (dark)
#}




contrast <- function(images){
  # We get all images corresponding to a digit

  n_images <- nrow(images)
  
  # We use a for loop to calculate the ink of each digit
  dark <- vector(mode="integer", length=n_images)
  jeff<- vector(mode="integer", length=n_images)
  for (i in 1:n_images) {
    #jeff[i] = contrast_digit(images[i, ])
    dark[i] = count_dark_pixels(contrast_digit(images[i, 1:392]))
  }
  return (dark)
}

plot_digit(contrast_digit(mnist$train$images[7,]))
count_dark_pixels(contrast_digit(mnist$train$images[1:4,]))


```

Getting data and running the functions 
```{r}
library(dplyr)

#get all ones and fours for training 
image1_train <- get_image_digit_train(1)
image4_train <- get_image_digit_train(4)

#get all ones and fours for testing 
image1_test <- get_image_digit_test(1)
image4_test <- get_image_digit_test(4)





#calculate features for training get the number of pixels for ones 
image_1_dark_train <- get_num_pixels(image1_train)
image_1_light_train <- contrast(image1_train)


#calculate features for training get size of the image for ones 
image_1_size_train <- get_size(image1_train)
image_1_size_light_train <- get_size_light(image1_train)

#calculate features for testing get the number of pixels for ones 
image_1_dark_test <- get_num_pixels(image1_test)
image_1_light_test <- contrast(image1_test)

#calculate features for testing get size of the image for ones 
image_1_size_test <- get_size(image1_test)
image_1_size_light_test <- get_size_light(image1_test)


#combine the tables to get our training and testing tables for ones training
combined_table_1_train <- data.frame(Size = image_1_size_train, Size_Contrast = image_1_size_light_train,   Dark = image_1_dark_train, Contrast = image_1_light_train,  Number = 1)
#testing
combined_table_1_test <- data.frame(Size = image_1_size_test, Size_Contrast = image_1_size_light_test,    Dark = image_1_dark_test, Contrast = image_1_light_test,  Number = 1)


#calculate features for training get the number of pixels for ones  

#image_4_dark_train <- get_num_pixels(image4_train)
#image_4_size_train <- get_size(image4_train)

#calculate features for training get the number of pixels for fours
image_4_dark_train <- get_num_pixels(image4_train)
image_4_light_train <- contrast(image4_train)

#calculate features for training get size of the image for fours 
image_4_size_train <- get_size(image4_train)
image_4_size_light_train <- get_size_light(image4_train)



#image_4_dark_test <- get_num_pixels(image4_test)
#image_4_size_test <- get_size(image4_test)


#calculate features for testing get the number of pixels for fours 
image_4_dark_test <- get_num_pixels(image4_test)
image_4_light_test <- contrast(image4_test)

#calculate features for testing get size of the image for fours 
image_4_size_test <- get_size(image4_test)
image_4_size_light_test <- get_size_light(image4_test)



#combined_table_4_train <- data.frame(Size = image_4_size_train, Dark = image_4_dark_train, Number = 4)
#combined_table_4_test <- data.frame(Size = image_4_size_test, Dark = image_4_dark_test, Number = 4)

#combine the tables to get our training and testing tables for ones training
combined_table_4_train <- data.frame(Size = image_4_size_train, Size_Contrast = image_4_size_light_train,  Dark = image_4_dark_train, Contrast = image_4_light_train,  Number = 4)
#testing
combined_table_4_test <- data.frame(Size = image_4_size_test, Size_Contrast = image_4_size_light_test, Dark = image_4_dark_test, Light = image_4_light_test,  Number = 4)



combined_train <- bind_rows(combined_table_1_train, combined_table_4_train)
combined_test <- bind_rows(combined_table_1_test, combined_table_4_test)

```


sampling our table 
```{r}

set.seed(123)
sampled_data_train <- combined_train %>% sample_n(800, replace = FALSE)
sampled_data_test <- combined_test %>% sample_n(200, replace = FALSE)
```


```{r}
#need separate paths for separate data sets. -Mitch
file_path_test <- "~/Mscs 341 F23/Project/Mitch, Evan, Adam/Data/sampled_data_test.csv"
file_path_train <- "~/Mscs 341 F23/Project/Mitch, Evan, Adam/Data/sampled_data_train.csv"

# Write the files -Mitch
write.csv(sampled_data_train, file = file_path_train)
write.csv(sampled_data_test, file = file_path_test)



```


